---
title: "Midterm"
author: "Teryl Schmidt"
date: "10/6/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Additional Pre Midterm 1 Practice Questions: Stat 324

#1
Jane’s dog will have a litter of 3 puppies equally likely to be a boy or girl, independent of the other.
Define the events:

2 * 2 * 2 = 8 total

MMM
MMF
MFF
MFM
FMF
FMM
FFM
FFF

A All the puppies are the same gender

```{r}
# P(# of favorable outcomes / Total # of outcomes)
# P(2 / 8)
# P(A) = P(MM) + P(FF) 
# = 2(1/2)^3 (each outcome is 1/2 (but double because can be all male or all female), doing 3 pulls)
# = 1/4
```

2/8 = 1/4

B There is at most one male puppy

```{r}
# P(# of favorable outcomes / Total # of outcomes)
# P(B) = P(FF) + P(FFF)
# = (1/2)^3 + 1/8
# = 3/8 + 1/8 = 4/8 = 1/2
```

4/8 = 1/2

C The litter includes a male and a female.

```{r}
# P(# of favorable outcomes / Total # of outcomes)
# P(C) = 1 - (P(MMM) + P(FFF))
# = 1 - (1/8 + 1/8)
# = 1 - 2/8 = 6/8
```

6/8 = 3/4

Which pairs of events, if any, are independent? Which pairs of events, if any, are mutually exclusive?

```{r}
# P(A|B) = P(A and B) / P(B)
# P(A|B) = (P(A) * P(B)) / P(B)
# if P(A) == P(A|B) then A and B are independent
# if don't equal then mutually exclusive
```

A and B are independent
P(A|B) = (1/8) / (1/2) = 1/4
P(A) = 1/4
since P(A|B) = P(A) = 1/4 or P(A) ∗ P(B) = (1/4) ∗ (1/2) = P(A and B) = 1/8 

B and C are independent
P(B|C) = (3/8) / (3/4) = 1/2
P(B) = 1/2
since P(B|C) = P(B) = 1/2 or P(B) * P(C) = (1/2) * (3/4) = P(B and C) = 3/8

A and C are not independent
P(A|C) = (0/8) / (3/4) = 0
P(C) = 3/4
since P(A|C) = P(C) = 3/4 or P(A) * P(C) = (1/4) * (3/4) = P(A and C) = 3/16

A and C are mutually exclusive since they do not contain any of the same outcomes. P(A and C) = 0.
Mutually exclusive events are extremely dependent.


```{r}
# Events are mutually exclusive if the occurrence of one event excludes the occurrence of the other(s). Mutually exclusive   events cannot happen at the same time. For example: when tossing a coin, the result can either be heads or tails but cannot be both.

#Events are independent if the occurrence of one event does not influence (and is not influenced by) the occurrence of the other(s). For example: when tossing two coins, the result of one flip does not affect the result of the other.
```


#2
An object with initial temperature T0 is placed in an environment with ambient temperature Ta.
According to Newton’s law of cooling the temperature T of the object is given by T = cT0 + (1 − c)Ta,
where c is a constant that depends on the physical properties of the object and on the elapsed time.
Assume that T0 has mean 25◦ C and standard deviation 2◦ C, and Ta has mean 5◦ C and standard
deviation 1◦ C. Find the mean and standard deviation of T at the time when c = 0.25 assuming T0 and
Ta are independent.

Mean <br/>
E(T) = E(cT0 + (1 − c)Ta) <br/>
= cE(T0) + (1 − c)E(Ta) <br/>
= 0.25(25) + (1 - 0.25)(5) <br/>
= 6.25 + 3.75 = 10 <br/>

Variance <br/>
Var(T) = Var(cT0 + (1 - c)Ta) <br/>
= c<sup>2</sup>Var(T0) + (1 - c)<sup>2</sup>Var(Ta) <br/>
= 0.25<sup>2</sup>Var(2<sup>2</sup>) + (1 - 0.25)<sup>2</sup>Var(1<sup>2</sup>) <br/>
= 0.25 + (0.5625)(1) = 0.8125 <br/>
= sqrt(0.8125) = 0.9013878 <br/>


#3
Suppose in Ms Green’s large introductory statistics class, the distribution of X=raw scores on midterm
1 is approximately Normally distributed with a mean of 17.2 points and a standard deviation of 3.8
points. The professor decides to scale the scores by multiplying the raw score by 4 and adding 10.

(a) Define the random variable Y = 4X + 10 to be the scaled score of a randomly selected student
from this class. Determine the shape, mean, and standard deviation of Y

Normal since linear transf of Normal, E(Y) = 4(E(X)) + 10 = 4(17.2) + 10 = 78.8,
SD(Y ) = |4|(SD(X)) = 4(3.8) = 15.2. <br/>


(b) What is the probability that a randomly selected student has a scaled test score of at least 90?

P(X >= 90) <br/>
1 - P(Z <= 90) <br/>
1 - P(Z <= 90 - 78.8 / 15.2) <br/>
1 - P(Z <= 0.7368421) <br/>
1 - 0.7704 = 0.2296 <br/>


(c) What is the probability that if we randomly select 2 students in the class, the sum of their scaled
test scores is at least 180 assuming their scores are independent?

S is Normal because sum of normal E(S) = E(X) + E(X) = 78.8 + 78.8 = 157.6 <br/>
V ar(S) = V ar(X) + V ar(X) = 2(15.2<sup>2</sup>) = 462.08 <br/>
Sd(S) = sqrt(462.08) = 21.5 <br/>
P(S ≥ 180) <br/>
= 1 − P(S ≤ 180) <br/> 
= 1 − P(Z ≤ (180 − 157.6 / 21.5) <br/>
= 1 − P(Z < 1.04186) = 0.1487 <br/>


#4
The diameter of hail hitting the ground during a storm is well approximated by a normal distribution
with a mean of 0.5 inch and a standard deviation of 0.1 inch.
For parts a and b, what is the probability that:

(a) A hailstone picked up at random will have a diameter greater than 0.71 inch?

P(X > 0.71) <br/>
1 - P(Z < 0.71) <br/>
1 - P(Z < 0.71 - 0.5 / 0.1) <br/>
1 - P(Z < 2.1) = 1 - 0.9821 = 0.179 <br/>

(b) Two hailstones picked up in a row will have diameters greater than 0.6 inch each? (Assume
independence of the two diameters)

P(X > 0.6) <br/>
1 - P(Z < 0.6) <br/>
1 - P(Z < 0.6 - 0.5 / 0.1) <br/>
1 - P(Z < 1) <br/>
1 - 0.8413 = 0.1587 <br/>

P(D1 > .6andD2 > .6) <br/>
= P(D1 > .6) ∗ P(D2 > .6)  <br/>
= (1 − (P(D1) < .6))<sup>2</sup> <br/> 
= (1 − 0.8413447)<sup>2</sup> <br/>
= 0.1586553<sup>2</sup> = 0.0251715 <br/> 
pnorm(.6,.5,.1) = 0.8413447 <br/>


(c) By the end of the storm, what proportion of the hailstones would have had diameters between 0.4
inch and 0.6 inch?

68% since within one SD from mean

P(0.4 < X < 0.6)
P(0.4) - P(0.6)
P(x - 0.5 / 0.1 < 0.6 - 0.5 / 0.1) - P(x - 0.5 / 0.1 < 0.4 - 0.5 / 0.1)
P(x - 0.5 / 0.1 < 1) - P(x - 0.5 < -1)
P( -1 < X < 1)
ztable - ztable
0.8413 - 0.8413 = 0 for probability


(d) If only 5% of hailstones have diameters above d, what is the value of d?

P(X > d) = 0.05 <br/>
P(X <= d) = 0.95 <br/>
P(Z <= d) = 0.95 look for 0.95 in the table <br/>
P(Z <= d) = 1.65 <br/>
1.65 = (d - 0.5 / 0.1) <br/> 
d = .1 ∗ 1.645 + .5 = 0.6645 <br/>

(e) If Estelle picks up 6 hailstones, what is the probability that exactly one of them has a diameter
above 0.55?

Let Y = number of hailstones with diameter above 0.55. <br/>
Y ∼Bin(6, p) where p = P(D > 0.55) = 1 − P(D ≤ 0.55) = 0.3085375. <br/> 
So Using Binomial Formula: <br/>
(6 choose 1) ∗ 0.309<sup>1</sup> ∗ 0.691<sup>5</sup> = 0.293 <br/>


#5
Suppose there are two three faced dies with number of points 1,2,3. One die is fair and the other one is
not. Let X denotes the number obtained when the fair die is rolled and Y denotes the number obtained
when the unfair die is rolled. Assume X and Y to be independent. Suppose, Y has probability mass
function (pmf) pY (y) = c ∗ y for y = 1, 2, 3.

(a) Write the pmf of X. Find E(X) and Var(X).

p(x) = 1/3 for x = 1, 2, 3 <br/>
E(X) = 1 ∗ (1/3) + 2 ∗ (1/3) + 3 ∗ (1/3) = 2 <br/>
Var(X) = (1 − 2)<sup>2</sup> ∗ (1/3) + (2 − 2)<sup>2</sup> ∗ (1/3) + (3 − 2)<sup>2</sup> ∗ (1/3) = .67 <br/>


(b) Find the constant c such that pY (y) is a proper pmf.

&#931;<sup>2</sup><sub>i=1</sub> c * y = 1 <br/>
i.e. c + 2c + 3c = 1 <br/>
i.e. c = 1/6 so pmf: <br/>

            Y  |  1  |  2  |  3  |
      P(Y = y) | 1/6 | 2/6 | 3/6 |


(c) Find E(Y) and Var(Y)

E(Y) = (1 ∗ 1/6) + (2 ∗ 2/6) + (3 ∗ 3/6) = 2.33
Var(Y) = (1 − 2.33)<sup>2</sup> ∗ 1/6 + (2 − 2.33)<sup>2</sup> ∗ 2/6 + (3 − 2.33)<sup>2</sup> ∗ 3/6 = 0.5555667


(d) Let Z denotes another RV denoting average of the no. of points obtained when each of two dies
rolled once, i.e. Z = (X + Y)/2. Find E(Z) and Var(Z). [You can use that E(Y) = 2.33 and
Var(Y) = 0.56]


E(Z) = E(X) + E(Y) / 2 <br/>
= (2 + 2.33) / 2 = 2.165. <br/>
Var(Z) = Var(X) + Var(Y) / 4 <br/>
= 0.67 + 0.56 / 4 <br/>
= 0.3075 <br/>
Note, Z is not Normally Distributed; neither X nor Y are <br/>


(e) Determine E(Y − 2X) and V ar(Y − 2X) What does Y - 2X mean in context?.

E(Y − 2X) <br/>
= E(Y) − 2 ∗ E(X) <br/> 
= 2.33 − 2 ∗ 2 <br/>
= −1.67; <br/>
Var(Y − 2X) <br/>
= Var(Y) + 4Var(X) <br/> 
= 0.56 + 4 ∗ 0.67 <br/>
= 3.24 <br/>
Using the fact that X and Y are independent. Y - 2X is the difference between the face rolled by the
Y dice and twice that rolled by the X.


#6
A patient visits her doctor with concerns about her blood pressure. If the systolic blood pressure
exceeds 150, the patient is considered to have high blood pressure and medication may be prescribed.
The problem is that there is considerable variation in a patient’s systolic blood pressure readings during
a given day.

(a) If a patient’s systolic readings during a given day have a normal distribution with a mean of 160
mm mercury and a standard deviation of 20 mm, what is the probability that a single measurement
will fail to detect that the patient has high blood pressure?

X ∼ N(160, 20<sup>2</sup>) <br/>
P(Fail to detect) = P(X < 150) <br/>
= P(Z < 150−160 / 20) <br/>
= P(Z < −0.5) = 0.3085. <br/>
There’s over a 30% chance of failing to detect patient has high blood pressure if
only a single measurement is taken. <br/>


(b) If five measurements are taken at various times during the day, what is the probability that the
average blood pressure reading will be less than 150 and hence fail to indicate that the patient has
a high blood pressure problem?

Let Y be the average blood pressure of the five measurements. Then, Y has a normal distribution with <br/>
µ = 160 and σ = 20/√5 = 8.944. <br/>
P(Y ≤ 150) = P(Z ≤ 150 − 160 / 8.944) <br/> 
= P(Z ≤ −1.12) = 0.1314 <br/>
Therefore, by using the average of five measurements, the chance of failing to detect the patient has <br/>
high blood pressure has been reduced from over 30% to about 13% <br/>


(c) How many measurements would be required so that the probability is at most 1% of failing to
detect that the patient has high blood pressure (by considering the average of the measures)?

Failing to detect means Y < 150, so we need to determine the sample size n such that P(Y¯ ≤ 150) ≤ 0.01 <br/>
Now, P(Y¯ ≤ 150) = P(Z ≤ 150 − 160 / 20/√n) <br/>
From the normal table, we can have P(Z ≤ −2.33) ≈ 0.01 <br/>
Therefore, 150−160 / 20/√n = −2.33 <br/>
Solving for n, yields n = 20<sup>2</sup>(2.33<sup>2</sup>) / 10<sup>2</sup> = 21.72 <br/>
round up to the closest integer will give 22 measurements. <br/>
So, it would require at least 22 measurements in order to achieve the goal of at most a 1% chance <br/>
of failing to detect high blood pressure (assuming patient has true mean blood pressure of 160 with sd = 20). <br/>


#7
What sample size is required in order that the standard deviation of X is at most

(a) 1/4 of the population standard deviation?

1/4 = σ /√n <br/>
σ = 1 | √n = 4 <br/>
SD(X) = σ /√n <br/>
when n = 16 <br/>


(b) 1/7 of the population standard deviation?

1/7 = σ /√n <br/>
σ = 1 | √n = 7 <br/>
SD(X) = σ /√n <br/>
when n = 49 <br/>


(c) 12% of the population standard deviation?

SD(X) = σ / √n = .12σ <br/>
when n = (1 / 0.12)<sup>2</sup> = 69.444 <br/>
so sample size of 70 to be sure it is less than 12% <br/>


#8
Daily carbon monoxide (CO) emission from a large production plant will be measured on 25 randomly
seleted weekdays. The production process is always being modified and the current mean value of
daily CO emissions µ is unknown. Data collected over several years confirm that, for each year, the
distribution of CO emissions is normal with a standard deviation of 0.8 tons.

(a) Suppose the sample mean is found to be x = 2.7 tons. Construct a 95% confidence interval for µ
assuming a known population sd and interpret it in context.

2.7 ± (0.025 in z-table = |-1.96|) <br/>
2.7 ± 1.96 ∗ 0.8 / √25 = (2.3864, 3.0136) <br/>
We are 95% confident the true mean amount of daily carbon monoxide emissions from a large production plant is between 2.39 and 3.01 (since 95% of the confidence intervaals we could have constructed would have captured it.)


(b) Suppose the scientist does not believe 0.8 is an accurate measure of the population sd and chooses
to instead use the sample standard deviation of sx = 1.3 tons. How would the scientist’s confidence
interval change?

It gets wider because S is larger and T is more spread out: <br/>
2.7 ± t(0.025,24) ∗ 1.3 / √25 <br/>
= 2.7 ± 2.064 ∗ 1.3 / √25 <br/>
= 2.7 ± 0.53664 <br/>
= (2.16336, 3.23664) <br/>




Bernoulli pi(1 - pi)

Binomial n * pi(1 - pi)

CLT




#2 on homework estimatior






