---
title: "Discussion9"
author: "Teryl Schmidt"
date: "11/6/2018"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#Ex 1: 
AA<-c(0.10, 0.12, 0.08, 0.14, 0.08, 0.11, 0.13, 0.09, 0.12, 0.15)
BB<-c(0.14, 0.13, 0.10, 0.19, 0.01, 0.28, 0.46, 0.24, 0.07, 0.10)

(xbar.AA<-mean(AA)); (s.AA<-sd(AA))
(xbar.BB<-mean(BB)); (s.BB<-sd(BB))

Ex1.combo<-c(AA, BB)
Ex1.labels<-c(rep("AA", length(AA)),rep("BB", length(BB)))
Ex1.df<-data.frame(Ex1.combo, Ex1.labels)

stripchart(Ex1.combo~Ex1.labels, method="stack")
qqnorm(AA)
qqnorm(BB)

na=length(AA)
nb=length(BB)
df<-(var(AA)/na+var(BB)/nb)^2/((var(AA)/na)^2/(na-1)+(var(BB)/nb)^2/(nb-1))
df
t.test(AA, BB, var.equal = FALSE, alternative = "two.sided")

#By Hand in R
(Ex1.tobs<-(xbar.AA-xbar.BB-0)/sqrt(s.AA^2/10+s.BB^2/10))
p.val<-2*pt(Ex1.tobs, df)
p.val

#Bootstrap for Two Sample Test
boottwo <- function(dat1, dat2, nboot) { 
  bootstat <- numeric(nboot)
  truediff <- mean(dat1) - mean(dat2)
  n1 <- length(dat1)
  n2 <- length(dat2) 
  for (i in 1:nboot) {
    samp1 <- sample(dat1, size = n1, replace = T) 
    samp2 <- sample(dat2, size = n2, replace = T) 
    bootmean1 <- mean(samp1)
    bootmean2 <- mean(samp2)
    bootvar1 <- var(samp1)
    bootvar2 <- var(samp2)
    bootstat[i] <- (bootmean1 - bootmean2 - truediff)/sqrt((bootvar1/n1) + (bootvar2/n2))
  }
  return(bootstat) 
}

n.boot<-10000

set.seed(1)
Ex1.boot<-boottwo(AA, BB, n.boot)
(m.low<-sum(Ex1.boot<= Ex1.tobs))  #521
(m.up<-sum(Ex1.boot >= Ex1.tobs))   #9479
(pval<-2*min(m.low, m.up)/n.boot)  #0.1042
```
(a) What problems might exist in analyzing this particular experiment?


(b) If we ignore these problems, what is the appropriate set of hypotheses if we wish to determine whether the mean proportion of e-Coli present is the same with the two antibiotics?


(c) What kind of tests could we perform and why? (Hint: think about the assumptions of each test.)


(d) Conduct the test(s) by hand at the 5% after computing the summary statistics (as applicable). Confirm your findings with R What are the p-values for those tests you identified in a? How do the conclusions of the tests compare?


(e) Suppose during the researcher’s literature review, they saw that other studies have proposed that antibiotic B has a proportion of 15% e-Coli after two weeks in agar plates prepared as they had. If instead of preparing agar plates with antibiotic B and collecting sample data, they used .15 as a true proportion of e-Coli with antibiotic B, how would their process change? Which experiment and analysis (original scenerio vs that proposed in this part) would be the most informative?



```{r}
#Question 2
AA.2<-c(5.42, 3.54, 2.14, 9.56, 17.16, 1.98, 1.30, 2.01, 2.23, 2.04, 2.10, 2.12, 2.09, 2.07)
BB.2<-c(0.31, 0.27, 0.93, 0.92, 0.18, 0.90, 0.64, 2.42, 2.87, 2.15, 2.53, 2.75, 2.34, 2.38)

(xbar.AA2<-mean(AA.2)); (s.AA2<-sd(AA.2))
(xbar.BB2<-mean(BB.2)); (s.BB2<-sd(BB.2))

Ex2.combo<-c(AA.2, BB.2)
Ex2.labels<-c(rep("AA2", length(AA.2)),rep("BB2", length(BB.2)))
Ex2.df<-data.frame(Ex2.combo, Ex2.labels)

stripchart(Ex2.combo~Ex2.labels, method="stack")
boxplot(Ex2.combo~Ex2.labels)
qqnorm(AA.2)
qqnorm(BB.2)

(Ex2.t<-(xbar.AA2-xbar.BB2-0)/sqrt(s.AA2^2/10+s.BB2^2/10))
set.seed(1)
Ex2.boot<-boottwo(AA.2, BB.2, n.boot)
hist(Ex2.boot)
(m.low2<-sum(Ex2.boot<= Ex2.t))  #9883
(m.up2<-sum(Ex2.boot >= Ex2.t))  #117
(pval2<-2*min(m.low2, m.up2)/n.boot)
```


(c) What kind of tests could we perform and why? (Hint: think about the assumptions of each test.)


(d) Conduct the test(s) by hand at the 5% after computing the summary statistics (as applicable). Confirm your findings with R What are the p-values for those tests you identified in a? How do the conclusions of the tests compare?



```{r}
#Exercise 3
AA.3<-c(2.19, 2.07, 2.28, 2.08, 2.12, 1.88, 1.99)
BB.3<-c(1.99, 1.90, 1.74, 1.83, 1.87, 1.79, 1.94, 1.71)

(xbar.AA3<-mean(AA.3)); (s.AA3<-sd(AA.3))
(xbar.BB3<-mean(BB.3)); (s.BB3<-sd(BB.3))

Ex3.combo<-c(AA.3, BB.3)
Ex3.labels<-c(rep("AA3", length(AA.3)),rep("BB3", length(BB.3)))
Ex3.df<-data.frame(Ex3.combo, Ex3.labels)

stripchart(Ex3.combo~Ex3.labels, method="stack")
boxplot(Ex3.combo~Ex3.labels)
qqnorm(AA.3)
qqnorm(BB.3)

#T.test Equal Variance
t.test(AA.3, BB.3, var.equal=TRUE, alternative="two.sided")

#T.test Non Equal Variance
t.test(AA.3, BB.3, var.equal=FALSE, alternative="two.sided")

#Bootstrap
(Ex3.tobs<-(xbar.AA3-xbar.BB3-0)/sqrt(s.AA3^2/7+s.BB3^2/8))
set.seed(1)
Ex3.boot<-boottwo(AA.3, BB.3, n.boot)
(m.low3<-sum(Ex3.boot<=Ex3.tobs))
(m.up3<-sum(Ex3.boot>=Ex3.tobs))
pval3<-min(m.low3,m.up3)/n.boot
pval3
```


(c) What kind of tests could we perform and why? (Hint: think about the assumptions of each test.)


(d) Conduct the test(s) by hand at the 5% after computing the summary statistics (as applicable). Confirm your findings with R What are the p-values for those tests you identified in a? How do the conclusions of the tests compare?



