---
title: "Lecture7_Code"
author: "Chelsey Green"
date: "6/27/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Proportion CI

```{r}
succ_samp<-rbinom(n=1000, size=100, prob=0.25)
succ_prop<-succ_samp/100
hist(succ_prop); abline(v=.25, col="red")

pop_succ<-c(rep(1, 25), rep(0, 75))
prop<-rep(0, n<-1000)
for (i in 1:n){
  samp<-sample(pop_succ, size=100, replace=TRUE)
  prop[i]<-mean(samp)
}
hist(prop, freq=FALSE, main="Sample Proportions, n=100, pi=0.25"); abline(v=.25, col="red")
```

Bootstrap CI
```{r}
Cotanine<-c(29, 30, 53, 75, 34, 21, 12, 58, 117, 119, 115, 134, 253, 289, 287)
qqnorm(Cotanine)
hist(Cotanine)
```

Bootstrap Code:
Here is one way to do this bootstrap using R:

Create a new function, bootstrap(x, n.boot), having two inputs:
   - x is a data vector
   - n.boot is the desired number of resamples from x
 It returns a vector of n.boot t-hat values.

```{r}
bootstrap = function(x, n.boot) {
  n = length(x)
  x.bar <- mean(x)
  t.hat <- numeric(n.boot) # create vector of length n.boot zeros
  for(i in 1:n.boot) {
    x.star <- sample(x, size=n, replace=TRUE)
    x.bar.star <- mean(x.star)
    s.star <- sd(x.star)
    t.hat[i] <- (x.bar.star - x.bar) / (s.star / sqrt(n))
}
  return(t.hat)
}
```
Use the bootstrap() function to get an approximate sampling
distribution of T for the Cotanine data.
```{r}
Cotanine.boot <- bootstrap(Cotanine, 1000)
# Plot the approximate sampling distribution.
hist(Cotanine.boot, freq=FALSE, xlab = "Bootstrap t-hat values",
     main = "Approximate Sampling Distribution of T")
```

Find quantiles for a 95% confidence interval.
```{r}
t.lower <- quantile(Cotanine.boot, probs=.025) # This is our t_{1 - alpha.2}.
t.upper <- quantile(Cotanine.boot, probs=.975) # This is our t_{alpha/2}.
t.lower
t.upper
# Make the interval.
n = length(Cotanine)
x.bar = mean(Cotanine)
s = sd(Cotanine)
ci.low  = x.bar - t.upper * s / sqrt(n) # This is our lower interval endpoint.
ci.high = x.bar - t.lower * s / sqrt(n) # This is our upper interval endpoint.
interval = c(ci.low, ci.high)
names(interval)=NULL
print(interval)

#Assuming T distribution
x.bar+qt(.975, df=14)*s/sqrt(n)
x.bar-qt(.975, df=14)*s/sqrt(n)

```

