---
title: "Discussion3"
author: "Teryl Schmidt"
date: "9/26/2018"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Problem 1

Airlines sometimes overbook flights. Suppose that for a plane with 100 seats, an airline takes 110
reservations. Define the variable X as the number of people who actually show up for a sold-out flight.
From past experience, the probability distribution of X is given in the following table:

```{r}
X = c(95, 96, 97, 98, 99, 100, 101, 102, 103, 104, 105, 106, 107, 108, 109, 110)
Px =  c(.05, .10, .12, .14, .24, .17, .06, .04, .03, .02, .01, .005, .005, .005, .0037, .0013)

```


(a) What is the probability that the airline can accommodate everyone who shows up for the flight?
P(x <= 100)
P(x = 95) + P(x = 96) + P(x = 97) + P(x = 98) + P(x = 99) + P(x = 100)
.05 + .10 + .12 + .14 + .24 + .17
= 0.82

(b) What is the probability that not all passengers can be accommodated?
P(x > 100)
P(x = 101) + P(x = 102) + P(x = 103) + P(x = 104) + P(x = 105) + P(x = 106) + P(x = 107) + P(x = 108) + P(x = 109) + P(x = 110)
.06 + .04 + .03 + .02 + .01 + .005 + .005 + .005 + .0037 + .0013
= 0.18

OR

1 - P(x <= 100)
1 - 0.82
= 0.18


(c) If you are trying to get a seat on such a flight and you are number 1 on the standby list, what is
the probability that you will be able to take the flight? What if you are number 3?
P(x <= 99)
P(x = 95) + P(x = 96) + P(x = 97) + P(x = 98) + P(x = 99)
= 0.65

P(x <= 97)
P(x = 95) + P(x = 96) + P(x = 97)
= 0.27


##Problem 2
Of all the weld failures in a certain assembly, 85% of them occur in the weld metal itself, and the
remaining 10% occur in the base metal and the cause is unknown in 5% of failures. A sample of 10
weld failures is examined.


(a) What is the probability that exactly five of them are weld metal failures?
b = Total # success
n = # of trials
π = probability of success

E(b) = n * π
var(b) = n * π * (1 - π)

b = 5
n = 10
π = 0.85

```{r}
(factorial(10) / (factorial(5) * factorial(10 - 5))) * (0.85^5) * (1 - 0.85)^(10 - 5)

dbinom(5, 10, 0.85)
```

= 0.00849



(b) What is the probability that fewer than 3 of them are base metal failures?
b <= 2
n = 10
π = 0.10

P(b = 0) + P(b = 1) + P(b = 2)

```{r}
dbinom(0, 10, 0.10) + dbinom(1, 10, 0.10) + dbinom(2, 10, 0.10)

# <=
pbinom(2, 10, 0.10)
```

0.9298092

(c) What is probability that none of them are unknown failures?
b = 0
n = 10
π = 0.05

```{r}
dbinom(0, 10, 0.05)
```

0.5987369


(d) Find the expected value for number of base metal failures in this sample.
E(b) = n * π
b = 
n = 10
π = 0.10

E(b) = 10 * 0.10
     = 1

(e) Find the standard deviation of the number of base metal failures assuming the population
distribution given.
var(b) = n * π * (1 - π)
var(b) = 10 * 0.10 * (1 - 0.10)
       = 0.90
sqrt(0.90) = 0.9486833


##Problem 3

You are adding Badger-themed bedazzle to your stripped overalls and are using both red and white
beads. You are interested in how the size of the bag of beads you purchase to select your beads from
changes the probablility of different samples of beads.
Consider taking a sample of size 3, where each bead is selected without replacement of previous
beads, from each of two populations:


(a) Small population were the bag of beads contains 7 White (W) beads and 3 Red (R) beads.
RRR RWR RRW WRR RWW WRW WWR WWW

RRR
3/10 * 2/9 * 1/8

WRR
7/10 * 3/9 * 2/8

RWR
3/10 * 7/9 * 2/8

RRW
3/10 * 2/9 * 7/8

(b) Large population were the bag contains 7000 W and 3000 R.

WRR
7000/10000 * 3000/9999 * 2999/9998

RWR
3000/10000 * 7000/9999 * 2999/9998

RRW
3000/10000 * 2999/9999 * 7000/9998


Compare with a sample of size 3 generated from a spinner having probability of 0.7 W and 0.3 R.
Small = 0.05833333
Large = 0.0629979


(a) Let A=[Exactly one White]. Calculate the probability of A for each case.

WRR
RWR
RRW

P(A) = P(WRR) + P(RWR) + P(RRW)
     = 0.175 + 0.189 + 0.189
     
(7/10 * 3/9 * 2/8) + (3/10 * 7/9 * 2/8) + (3/10 * 2/9 * 7/8) = 0.175

(7000/10000 * 3000/9999 * 2999/9998) + (3000/10000 * 7000/9999 * 2999/9998) + (3000/10000 * 2999/9999 * 7000/9998)  = 0.18899
 
(0.70 * 0.30 * 0.30) + (0.30 * 0.70 * 0.30) + (0.30 * 0.30 * 0.70)  = 0.189

0.553

(c) Let B=[At least one White]. Calculate the probability of B for each case.

P(b) + P(b`) = 1
1 - P(b`) = p(b)

P(b`) = no white
RRR

P(b) = 1 - P(RRR)
P(b) = 1 - 0.0083333
     = 0.991666


(d) Notice, that the spinner model is equivalent to sampling with replacement from either of the two
finite populations. The spinner model is a classical representation of a device with no memory, so
that the outcome of the current trial is independent of all previous trials. When drawing from
a population that is large relative to the sample size, the calculated conditional probability for
the second and third beads is very similar to the probability of each outcome calculated from the
spinner (an independent model). So, as long as our population is large relative to our sample,
the events are independent "enough" that sampling without replacement does not change the
probability of an event.





