---
title: "Lect1-Intro"
author: "Chelsey Green"
date: "6/13/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval=TRUE)
```

A certain reaction was run several times using each of two catalysts, A and B. The catalysts were supposed to control the yield of an undesirable side product. Results, in units of ounces yield, for 4 runs of catalyst A and 6 runs of catalyst B are as follows:

CATALYST A: 

4.4 3.4 2.6 3.8 

CATALYST B: 

3.4 1.1 2.9 5.5 6.4 5.0 


ANSWER THE FOLLOWING QUESTIONS:


1. What type of data is this? 

2. Does there appear to be a (meaningful) difference between the two catalysts? Why? 

3. What would help you answer question 2 more confidently?  

4. What is the range of values for each catalyst?

5. What is the median value for each catalyst? 

6. What is the mean value for each catalyst? 

7. What is the standard deviation for each catalyst?


Entering the data
```{r}
Yield_A<-c(4.4, 3.4, 2.6, 3.8)
Yield_A
Yield_B<-c(3.4, 1.1, 2.9, 5.5, 6.4, 5.0)
Yield_B

#You do not need to combine your data into a dataframe, but this can be a useful form for seeing the relationship between variables
all_Yields<-c(Yield_A, Yield_B)
all_Yields
Catalyst<-c(rep("A", times=4), rep("B", times=6))
Catalyst

Yield_data<-data.frame(Catalyst, all_Yields)
View(Yield_data)
str(Yield_data)
```

Graphing the data using dotplots (since small numeric data sets)
```{r}
?stripchart   #Gives me info on the stripchart function

stripchart(all_Yields)  #dot plot of all yields

stripchart(Yield_A)  #dot plot of Catalyst A yields
stripchart(Yield_B)  #dot plot of Catalyst B yields

stripchart(all_Yields~Catalyst) #dot plot of all yields broken out by catalyst

stripchart(all_Yields~Catalyst, data=Yield_data, xlab="Yield by Catalyst", method="stack")   
#Update label on x axis to be more descriptive than variable name
```
Estimate some Descriptive Statistics (range, mean, median, sd) for each of the Catalyst yields before calculating them below.









Calculating some Descriptive Statistics Using R
```{r}
range(Yield_A)
range_A<-range(Yield_A)
range_A
range_A[2]-range_A[1]
median(Yield_A)
mean(Yield_A)
sd(Yield_A)  #sample SD

#Calculate SD by hand?

range_B<-range(Yield_B)
range_B[2]-range_B[1]
median(Yield_B)
mean(Yield_B)
sd(Yield_B)

```

Suped-up graphing
```{r}
#install.packages("ggplot2")
require(ggplot2)

ggplot(data=Yield_data, aes(x=Catalyst, y=all_Yields, color=Catalyst))+
  geom_point()+
  theme_bw()+
  ylab("Yields")
```