---
title: "HW1"
author: "Teryl Schmidt"
date: "9/15/2018"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Problem 1
If you wanted to estimate the mean height of all the students at UW Madison, which one of the following
sampling strategies would be best? Why? Note that none of the methods are true simple random samples.

(a) Measure the heights of 100 students found in the gym during basketball intramurals.
(b) Measure the heights of the engineering majors.
<font color="red"> 
(c) Measure the heights of the students selected by choosing the first name on each page of a list of students
enrolled that semester.
</font>


## Problem 2
A zoologist collected wild lizards in the Southwestern United States. Thirty lizards from the genus Phrynosoma
were placed on a treatmill and their speed measured. The recorded speeds (meters/second) (the
fastest time to run a half meter) for the thirty lizards are summarized in the relative histogram below. (Data
Courtesy of K. Bonine *)

(a) Is the percent of lizards with recorded speed below 1.25 closest to: <font color="red"> 25% </font>, 50%, or 75%?
(b) In which interval are there more speeds recorded: <font color="red"> 1.5-1.75 </font> or 2-2.5?
(c) About how many lizards had recorded speeds above 1 meters/second?
<font color="green"> 

~0.04 + ~0.04 = 0.08 below 1 m/s

~0.92 above 1 m/s

92% of 30 lizards is 27.6 lizards

<font color="red"> ~28 lizards have recorded speeds above 1 m/s </font>

</font>

##Problem 3
In a sample of 30 men, the mean height was 179 cm with standard deviation of 6 cm. In a sample of 25
women, the mean height was 163 cm with standard deviation of 6 cm. If both samples were combined into
one larger group...

(a) What is the mean height for the combined group?

<font color="red"> ~157.4167 </font> 
```{r}
# (female count * female mean) + (male count * male mean)
# -------------------------------------------------------
#              (feamle count + male count)

((25 * 163) + (30 * 179)) / (25 + 35)
```

(b) The standard deviation for the combined group would be

    i. Less than 6 cm
    <font color="red"> 
    ii. Greater than 6 cm
    </font>
    iii. Equal to 6 cm 
    iv. Can not tell from the information given


```{r}
# (male count * ((male SD)^2 + (male mean - combined mean)) + (female count * ((female SD)^2 + (feamle mean - combined mean))
# ---------------------------------------------------------------------------------------------------------------
#                                             (feamle count + male count)

# sqare root answer

((30 * (6^2 + (179 - 157.4167))) + (25 * (6^2 + (163 - 157.4167)))) / (25 + 30)

sqrt(50.31057)
```

##Problem 4
After manufacture, computer disks are tested for errors. The table below gives the number of errors detected
on a random sample of 100 disks.

(a) Construct a frequency histogram of the information by creating bins at [-.5, .5), [.5, 1.5), etc.

```{r}
Defects = c(0,1,2,3,4)
Frequency = c(rep(0, 41), rep(1, 31), rep(2, 15), rep(3, 8), rep(4, 5))
hist(Frequency, main="Frequency of Defects of Disk's", xlab="Defects", breaks = seq(-.5, 4.5, 1))

```

(b) What is the shape of the histogram for the number of defects observed in this sample?

    <font color="red"> Right Skewed </font>

(c) Calculate the mean and median number of errors detected on the 100 disks. How do these values
    compare and is that consistent with what we would guess based on the shape?

```{r}
mean(Frequency)
```
```{r}
median(Frequency)
```

  <font color="red"> The mean is more than the median so the histograph is skewed to the right. </font>

(d) Calculate the sample standard deviation with your calculator and R. Are the values consistent between
    the two methods? Explain what this value means in the context of the problem.
    
```{r}
sd(Frequency)

mean = mean(Frequency)

for (value in Frequency) {
  value = (value - mean)^2
}

sqrt(mean(Frequency))
```
    
  <font color="red"> Not the exact same number due to rounding. This value means how far or an average value is from the mean.     </font>    

(e) Calculate the first and third quartiles and IQR by hand and with R. Are the values consistent betweeen
    the two methods? Explain what the three values mean in the context of the problem.

```{r}
#R IQR
IQR(Frequency)
quantile(Frequency, c(0.25, 0.75))

#Hand IQR
IQR = quantile(Frequency, prob=0.75)-quantile(Frequency, prob=0.25)
IQR
```

  <font color="red"> The IQR is how spread out the middle values are. The spread goes from 0 to 2, which means a lot of the        data is around 0 to 2, and there are not many data points from 2 to 4. </font>

(f) What proportion of the computer disks had a number of errors greater than the mean number of errors?

```{r}
# Mean = ~1
# Add up defects from 2 to 4, then divide by total

(15 + 8 + 5) / 100
```

  <font color="red"> ~28% </font>

(g) What range of values for this sample data are not considered outliers using the [Q1-1.5IQR, Q3+1.5IQR]
    designation (using the IQR you calculated by hand)?.

```{r}
# Q1 - 1.5IQR
0 - (1.5 * IQR)

# Q3 +1.5IQR
2 + (1.5 * IQR)
```
  <font color="red"> Outliar range = x < -3 , 5 < x </font>

(h) Sketch a boxplot of the data by hand (using the relevant values you calculated by hand).

```{r}
boxplot(Frequency, horizontal = T)
```

(i) Compare and contrast (briefly) the information about the data given by the histogram in part a and
the boxplot in part h.

  <font color="red"> A Boxplot shows the minimum value, the first and third quartile, and the mean value. The histogram displays the frequencies of values.  </font>
  

##Problem 5
Physical education researchers interested in the development of the overarm throw measured the horizontal
velocity of a thrown ball at the time of release. The results for first-grade children (in feet/sec) (courtesy of
L. Halverson and M. Roberton*) are:
Males: 54.2, 39.6, 52.3, 48.4, 35.9, 30.4, 25.2, 45.4, 48.9, 48.9, 45.8, 44.0, 52.5, 48.3, 59.9, 51.7, 38.6, 39.1,
49.9, 38.3
Females: 30.3, 43.0, 25.7, 26.7, 27.3, 31.9, 53.7, 32.9, 19.4, 23.7, 23.3, 23.3, 37.8, 39.5, 33.5, 30.4, 28.5
3

(a) Use R to create a histogram for the males and a histogram for the females (any kind of histogram that
you want). Adjust the x axis scale so the two groups are more easily compared.

```{r}
Males = c(54.2, 39.6, 52.3, 48.4, 35.9, 30.4, 25.2, 45.4, 48.9, 48.9, 45.8, 44.0, 52.5, 48.3, 59.9, 51.7, 38.6, 39.1, 49.9, 38.3)
Females = c(30.3, 43.0, 25.7, 26.7, 27.3, 31.9, 53.7, 32.9, 19.4, 23.7, 23.3, 23.3, 37.8, 39.5, 33.5, 30.4, 28.53)

hist(Males, main = "Male Velocity", xlab = "velocity",col = "blue", xlim = c(15,60), ylim = c(0,7))
hist(Females, main = "Female Velocity", xlab = "velocity",col = "pink", xlim = c(15,60), ylim = c(0,7))
```

(b) Compare the shape of the throws from the male and female students observed in this sample.

<font color="red"> Females is right skewed. Males is left skewed. </font>

(c) Compute and compare the mean and median throw velocities observed in the male and famale students
    across gender.

```{r}
mean(Males)
median(Males)

mean(Females)
median((Females))
```

<font color="red"> The mean and median of the males data is higher. </font>

(d) Compute and compare the standard deviation in throw velocities observed in the male and famale
    students.

```{r}
sd(Males)

sd(Females)
```

<font color="red"> The standard deviation of both the data sets are about the same. </font>

(e) Use R (or by hand if you prefer) to help you create a boxplots of the two sets so they are easily
    comparable.

```{r}
data = range(Males,Females)
boxplot(Males, Females, horizontal = T)
```

(f) Which, if any values were identified as outliers? Would this value have been identified as an outlier if
it were thrown by the opposite gender?

<font color="red"> 
Female = 53.7 Male = 25.2 

These values are the biggest outliars in the data. If the female and male outliars were to swap, then they would not be outliars.
</font>

##Problem 6
There are 12 numbers on a list, and the mean is 24. The smallest number on the list is changed from 11.9
to 1.19.

(a) Is it possible to determine the direction in which (increase/decrease) the mean changes? Or how much
    the mean changes? If so, by how much does it change? If not, why not?

```{r}
# Mean + ((after - before) / count)
24 + (((1.19 - 11.9) / 12))
```

  <font color="red"> The mean will decrease by 0.8925, making the new mean 23.1075 </font>

(b) Is it possible to determine the direction in which the median changes? Or how much the median
    changes? If so, by how much does it change? If not, why not?

  <font color = "red"> Yes, since the smallest number had been decreased it remained the smallest number. The smallest number in   the dataset does not effect the median so the median didn't change. </font>

(c) Is it possible to determine the direction in which the standard deviation changes? Or how much the
standard deviation changes? If so, by how much does it change? If not, why not?

```{r}
# ((orig val)^2 + (new val)^2 + count) * ((orig mean)^2 - (new mean)^2)) * ((1 / count - 1))

((-11.9^2 + 1.19^2) + (12 * (24^2 - 23.1075^2))) * (1 / (12 - 1))
```

  <font color = "red"> No, we cannot determine how much the standard deviation will change. But we can determine the square   of   the standard deviation will increase by 33.12067. </font>


##Problem 7
The UW Statisics Department is trying to determine what day of the week to hold their annual fall festival.
Assume that the weekdays, Monday through Friday, are equally likely and that each weekend day, Saturday
and Sunday, is three times as likely as a weekday to be selected.

(a) Assign probabilities to the seven outcomes.

<font color="red"> 
Monday = 1

Teusday = 1

Wednesday = 1

Thursday = 1

Friday = 1

Saturday = 3

Sunday = 3
</font>

(b) Find the probability a weekday will be selected.

  <font color="red"> Weekday = 5 / 11 </font>


##Problem 8
(*)Suppose you are eating at a pizza parlor with two friends. You have agreed to the following rule to decide
who will pay the bill. Each person will toss a coin. The person who gets a result that is different from the
other two will pay the bill. If all three tosses yield the same result, the bill will be shared by all. (It may be
helpful to list the outcomes in the sample space)

(a) Find the probability that only you have to pay.

<font color="red"> 
HHH, HHT, HTH, THH, HTT, THT, TTH, TTT 

THH, HTT = I pay is 1/4
</font>

(b) Find the probability that all three will share.

<font color="red"> HHH, TTT = share bill is 1/4 </font>

##Problem 9
(*) The following frequency table shows the classification of 58 landfills in a state according to their concentration
of the three hazardous chemicals arsenic, barium, and mercury.
If a landfill is selected at random, find the probability that it has:

(a) A high concentration of barium.

<font color="red"> 
The probability that its concentration is high in Barium is 16 / 58
</font>

```{r}
# s = 5 + 11
# m = 58
16 / 58
```

(b) A high concentration of mercury and low concentrations of both arsenic and barium.

<font color="red"> 
The probability that its concentration is high in Mercury and low in Arsenis and barium is 10 / 58
</font>

```{r}
# s = 10
# m = 58

10 / 58
```

(c) A high concentration of any one of the chemicals and low concentrations of the other two.

<font color="red"> 
The probability that its concentration is high in one chemical and low in the other two is 27 / 58
</font>

```{r}
# s = 9 + 8 + 10
# m = 58

27 / 58
```

(d) A high concentration of Mercury given it has a high concentration of Arsenic. Is having high concentrations
of Mercury and high concentrations of Arsenic independent in this set of landfills?

<font color="red"> 
The probability that its concentration is high in Mercury given that it is a high concentration of Arsenic is 0.3333
In this sample, having high concentrations of Mercury and Arsenic is dependent.
</font>

```{r}
# P(High Mercury | High Arsenic) = P(High Mercury and Arsenic) / P(High Arsenic) = 0.3333
#                                =        (1 + 5)              /  (1 + 3 + 5 + 9)

# Check for dependence
# is P(High Mercury) == P(High Mercury | High Arsenic)
#       (20 / 58)    ==  (6 / 18)
#       0.3448276    =/=  0.33333

# P(High Mercury | High Arsenic)
(1 + 5) / (1 + 3 + 5 +9)

# P (High Mercury)
((5 + 5 + 10) / 58)
```

(e) A high concentration of Barium given it has a low concentration of Mercury. Is having high concentrations
of Barium and low concentrations of Mercury independent in this set of landfills?

<font color="red"> 
The probability that its concentration is high in Barium given that it is a low concentration of Mercury is 0.2894737
In this sample having high Barium and low Arsenic is dependent.
</font>

```{r}
# P(High Barium | Low Mercury) = P(High Barium and Low Mercury) / P(Low Mercury) = 0.2894737
#                              =          (3 + 8)               /   (11 + 27)

# Check for dependence
# is P(High Barium) == P(High Barium | Low Mercury)
#       (16 / 58)   ==  (38 / 58)
#       0.2758621   =/=  0.2894737

# P(High Barium | Low Mercury)
(3 + 8) / (11 + 27)

# P(High Barium)
((1 + 3 + 4 + 8) / 58)
```





