---
title: "HW3"
author: "Teryl Schmidt | tschmidt6@wisc.edu | 9072604920"
date: "9/27/2018"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


#Problem 1
A chemical supply company ships a certain solvent in 10-gallon drums. Let X represent the number of
drums ordered by a randomly chosen customer. Assume X has the following probability mass function
(pmf). The mean and variance of X is : µX = 2.3 and σ<sup>2</sup><sub>x</sub> = 1.81:

       X   |  1  |  2  |  3  |  4  |  5  |
    p(X=x) | 0.4 | 0.2 | 0.2 | 0.1 | 0.1 |

(a) Find P(X ≤ 2) and describe what it means in the context of the problem.

<font color="green"> P(X <= 2) = P(X = 1) + P(X = 2) = 0.4 + 0.2 </font>

<font color="red"> P(X <= 2) = 0.6 </font>


(b) Let Y be the number of gallons ordered, so Y = 10X. Find the probability mass function (pmf) of Y.

<font color="red"> define new table (second table) </font>

           X    |   1  |   2  |   3  |   4  |   5  |
        Y = 10x | 10*1 | 10*2 | 10*3 | 10*4 | 10*5 |
        p(X=x)  |  0.4 |  0.2 |  0.2 |  0.1 |  0.1 |
    

        Y = 10x |  10 |  20 |  30 |  40 |  50 |
        p(X=x)  | 0.4 | 0.2 | 0.2 | 0.1 | 0.1 |


(c) Find the mean number of gallons ordered µ<sub>Y</sub> using the pmf from part (b) and a second time using
the expectectation of linear combination formula shown in class.

<font color="green"> 
µ<sub>Y</sub> = E(Y) <br/>
= E(10X) <br/>
= 10E(X) <br/>
= 10 * µ<sub>X</sub> <br/>
= 10 * 2.3 <br/>
= <font color="red"> 23 </font> <br/>
</font>


(d) Find the standard deviation of the number of gallons ordered σ<sub>Y</sub> using the pmf from part (b) and
a second time using the standard deviation of linear combinationformulas shown in class.

<font color="green"> σ<sub>Y</sub> = &#8730;<span style="text-decoration:overline"> σ</span><sup>2</sup><sub>Y</sub> <br/>
= &#8730;<span style="text-decoration:overline"> 181 </span> <br/>
= <font color="red"> 13.454 </font>
</font>


(e) What possible values of Y are within two standard deviations of the mean value (that is, in the
interval from (µ − 2σ, µ + 2σ))? What is the probability that the observed value of Y is within
two standard deviations of the mean value?

<font color="green"> 
(23 - 13.454 , 23 + 13.454) <br/>
(9.546 , 36.454) <br/>
P(X = 1) + P(X = 2) + P(X = 3) <br/>
= <font color="red"> 0.8 </font>
</font>


#Problem 2
Consider a large population which has true mean µ and true standard deviation σ. We take a sample of
size 3 from this population, thinking of the sample as the RVs X<sub>1</sub>, X<sub>2</sub>, X<sub>3</sub> where X<sub>i</sub> can be considered iid. We are interested in estimating µ.


(a) Consider the estimator µ&#770;<sub>1</sub> = X<sub>1</sub> + X<sub>2</sub> − X<sub>3</sub>. Is this estimator unbiased? Calculate the bias.

<font color="green"> 
E(µ&#770;) = E(X<sub>1</sub> + X<sub>2</sub> − X<sub>3</sub>) <br/>
= E(X<sub>1</sub>) + E(X<sub>2</sub>) − E(X<sub>3</sub>) <br/>
= µ + µ - µ <br/>
= µ <br/>
<font color="red"> Therefore µ&#770;<sub>1</sub> = X<sub>1</sub> + X<sub>2</sub> − X<sub>3</sub> is an unbiased estimator for µ </font>
</font>

(b) Find the variance of µ&#770;<sub>1</sub>.

<font color="green"> 
V(µ&#770;) = V(X<sub>1</sub> + X<sub>2</sub> − X<sub>3</sub>) <br/>
= V(X<sub>1</sub>) + V(X<sub>2</sub>) − V(X<sub>3</sub>) <br/>
= σ<sup>2</sup> + σ<sup>2</sup> + σ<sup>2</sup> <br/>
= 3σ<sup>2</sup> <br/>
= <font color="red"> The variance of µ&#770;<sub>1</sub> is 3σ<sup>2</sup> </font>
</font>

(c) When estimating µ, explain why someone would prefer the estimator 
X&#772; = (X<sub>1</sub>+X<sub>2</sub>+X<sub>3</sub> / 3) over µ&#770;<sub>1</sub>

<font color="green"> 
µ&#770;<sub>2</sub> = (X<sub>1</sub>+X<sub>2</sub>+X<sub>3</sub> / 3) <br/>
E(X&#772;) = µ <br/>
E(µ&#770;) = E(X<sub>1</sub>+X<sub>2</sub>+X<sub>3</sub> / 3) <br/>
= \(\frac{1}{3}\) [E(X<sub>1</sub>) + E(X<sub>2</sub>) + E(X<sub>3</sub>)] <br/>
= \(\frac{1}{3}\) [µ + µ + µ] <br/>
= \(\frac{3µ}{3}\) <br/>
= µ <br/> <font color="red">unbiased</font>

V(X&#772;) = V(X<sub>1</sub> + X<sub>2</sub> − X<sub>3</sub> / 3) <br/>
= V(X<sub>1</sub>) + V(X<sub>2</sub>) − V(X<sub>3</sub>) <br/>
= σ<sup>2</sup> + σ<sup>2</sup> + σ<sup>2</sup> / 3<br/>
= 3σ<sup>2</sup> / 3<br/>
= σ<sup>2</sup> <font color="red">smaller variance</font>

<font color="red"> Therefore µ&#770;<sub>2</sub> = (X<sub>1</sub>+X<sub>2</sub>+X<sub>3</sub> / 3) is an unbiased estimator for µ. And we choose X&#772; over µ&#770;<sub>1</sub> because X&#772; has a smaller variance (3σ<sup>2</sup> / 3) = σ<sup>2</sup>, as oppose to 3σ<sup>2</sup> </font>
</font>


(d) Now consider the estimator µ&#770;<sub>2</sub> = (X<sub>1</sub>+X<sub>2</sub>+X<sub>3</sub> / 2). Is this estimator unbiased? Calculate the bias.

<font color="green"> 
µ&#770;<sub>2</sub> = (X<sub>1</sub>+X<sub>2</sub>+X<sub>3</sub> / 2) <br/>
E(X&#772;) = µ <br/>
E(µ&#770;) = E(X<sub>1</sub>+X<sub>2</sub>+X<sub>3</sub> / 2) <br/>
= \(\frac{1}{2}\) [E(X<sub>1</sub>) + E(X<sub>2</sub>) + E(X<sub>3</sub>)] <br/>
= \(\frac{1}{2}\) [µ + µ + µ] <br/>
= \(\frac{3µ}{2}\)  =/= µ <br/>
<font color="red"> Therefore µ&#770;<sub>2</sub> = (X<sub>1</sub>+X<sub>2</sub>+X<sub>3</sub> / 2) is a biased estimator for µ </font>
</font>


(e) Compute the MSE for µ&#770;<sub>2</sub>.

<font color="green"> 
µ&#770;<sub>2</sub> = (X<sub>1</sub>+X<sub>2</sub>+X<sub>3</sub> / 2) <br/>
Var(µ&#770;) = Var(X<sub>1</sub>+X<sub>2</sub>+X<sub>3</sub> / 2) <br/>
= \(\frac{1}{4}\) [Var(X<sub>1</sub>) + Var(X<sub>2</sub>) + Var(X<sub>3</sub>)] <br/>
= \(\frac{1}{4}\) [σ<sup>2</sup> + σ<sup>2</sup> + σ<sup>2</sup>] <br/>
= \(\frac{3σ^2}{4}\) <br/>
MSE(µ&#770;) = Var(µ&#770;) + [Biar(µ&#770;)]<sup>2</sup> <br/>
= MSE(µ&#770;) = \(\frac{3σ^2}{4}\) + (\(\frac{µ}{4}\))<sup>2</sup> <br/>
= MSE(µ&#770;) = \(\frac{3σ^2}{4}\) + (\(\frac{µ^2}{4}\)) <br/>
= MSE(µ&#770;) = <font color="red"> \(\frac{3σ^2 + µ^2}{4}\) </font> <br/>
</font>


#Problem 3
Let F be an RV that represents the operating temperature in Fahrenheit of one instance of a manufacturing process, and assume F ∼ N(100, Var(F) = 5<sup>2</sup>). Let C be an RV that represents the same process, but measured in Celsius. 

Fahrenheit can be converted to Celsius using C = \(\frac{5}{9}\) (F − 32). Using normal table in Canvas, solve for the following (You can also check your answers using R):


(a) Find the probability that one randomly selected instance of the process will have operating
temperature greater than 98.6 Fahrenheit.

<font color="green"> 
P(X > 98.6) <br/>
P(X - 100 / 5 > 98.6 - 100 / 5) <br/>
1 - P(Z <= -0.28) <br/>
1 - 0.3897 <br/>
= <font color="red"> 0.6103 </font>
</font>


(b) Find the distribution of C. (Hint: C ∼?(?, ?))

<font color="green"> 
distribution of will be normal with the mean <br/>
µ<sub>c</sub> = E(\(\frac{5}{9}\)(X - 32)) <br/>
= \(\frac{5}{9}\)E(E(X) - 32) <br/>
= \(\frac{5}{9}\)E(100 - 32)  <br/>
= <font color="red"> 340 / 9 = 37.77778 </font>


and variance <br/>
σ<sup>2</sup><sub>c</sub> = (5/9)<sup>2</sup>[Var(X) + 0] <br/>
= (5/9)<sup>2</sup>[5<sup>2</sup> + 0] <br/>
= <font color="red"> 7.716 </font>
</font>

<font color="red"> C ~N(37.77, 7.716) </font>

(c) Find the probability that one randomly selected instance of the process will have operating
temperature below 32 Celsius.

<font color="green"> 
P(X < 32) <br/>
= P(32 - 37.77778 / &#8730;7.716) <br/>
= P(Z < -2.0799) <br/>
= <font color="red"> 0.0188 </font><br/>
</font>

(d) Above what temperature (in Celsius) is the top 10% of operating temperatures?

<font color="green"> 
P(X > a) = 0.90 <br/>
P(x - 37.77 / &#8730;7.716 > a - 37.77 / &#8730;7.716) = 0.1 <br/>
(a - 37.77 / &#8730;7.716) = 0.90 <br/>
= 1.28 (table) <br/>
a = 37.77 + 1.28 + &#8730;7.716 <br/>
a = <font color="red"> 41.34 </font>
</font>


(e) Find the probability in a sample of 6 instances, more than 4 instances have operating temperature above 32 Celsius.(Assuming observations in the sample are independent)

<font color="green"> 
P(X > 32) <br/>
= 1 - P(32 - 37.77778 / &#8730;7.716) <br/>
= 1 - P(Z > -2.0799) <br/>
= 1 - 0.0188 <br/>
= <font color="red"> 0.9812 </font><br/>


P(X > 32) = 0.9812 <br/>
B = # of instances samples > 32 <br/>
X ~Bin(6, 0.9812) <br/>
P(X > 4) = P(X = 5) + P(X = 6) <br/>
= ((<sup>6</sup><sub>5</sub>) * (0.98)<sup>5</sup> * (1 - 0.98) + (<sup>6</sup><sub>6</sub>) * (0.98)<sup>6</sup> <br/>
= <font color="red"> 0.9943 </font> <br/>
</font>

(f) Find the distribution of X&#772; for n=6, then find the probability that the average operating temperature
in a sample of 6 instances is above 32 Celsius.

<font color="green"> 
σ<sup>2</sup> / n <br/>
= 7.716 / 6 = 1.29 <br/>
X&#772; ~N(37.77, 1.29) <br/>
P(X&#772; > 32) <br/>
= P(X&#772; - 37.77 / &#8730;1.29 > 32 - 37.77 / &#8730;1.29) <br/>
= 1 - (Z > -5.09) <br/>
= <font color="red"> 0.99999 or 1 </font>
</font>

#Problem 4
Retail stores experience their heaviest returns on December 26<sup>th</sup> and December 27<sup>th</sup> each year. The
distribution for the Number of Items Returned (X) for Hilldale Macy’s on those days last year is given
in the table below. It has mean: µ = 2.61 and variance σ<sup>2</sup> = 1.80 Assume the probability distribution
also holds for this year.
Number of Items Probability

    1 | 0.25
    2 | 0.28
    3 | 0.20
    4 | 0.17
    5 | 0.08
    6 | 0.02
(a) In this year, a random sample of size 45 returns is selected for review. Describe the sampling
distribution of the sample mean (shape, center, and spread).

<font color="green"> 
µ<sub>X&#772;</sub> = µ = 2.61 <br/>
σ<sub>X&#772;</sub> = σ / &#8730;n = &#8730;(1.80 / 45) = 0.20 <br/>
<font color="red"> The sample follows a normal distribution <br/> 
X&#772; ~N(2.61, 0.20) <br/>
</font> </font>

(b) What is the probability that the sample mean will be greater than 2.9 items?

<font color="green">
P(X&#772; > 2.9) = P(Z > 2.9 - 2.61 / 0.20) <br/>
P(Z > 1.45) = 0.5 - 0.4265 = <font color="red"> 0.0735 </font>
</font>

(c) Find an upper bound b such that the total number of items returned by 45 customers will be less
than b with probability 0.95.

<font color="green">
For 95% the Z value is 1.645 <br/>
Therefore the upperbound b is: <br/>
b = µ + Zσ = 2.61 + 1.645 * 0.20 = 2.939 <br/>
(Z <= x - 2.61 / 0.20) = 0.95 <br/>
x = 2.939 <br/>
b = 45 * 2.939 = <font color="red"> 132 </font>
</font>

#Problem 5
We will be exploring the difference between using the standard deviation formula: s1 = &#8730;<span style="text-decoration:overline">(X−X&#772;)</span><sup>2</sup> / &#8730;<span style="text-decoration:overline">(n - 1)</span>

s2 = &#8730;<span style="text-decoration:overline">(X−X&#772;)</span><sup>2</sup> / &#8730;<span style="text-decoration:overline">n</span> through a simulation.

In the code below, I have defined a population of values, named pop1. I have also written a function
sample.sd to compute the sample standard deviation on a set of numbers passed in.


(a) Copy and paste the entire chuck of code from set.seed(1) (so that we’re ll using the same data)
through par(mfrow=c(1,1)) (so we reset the graphics pane). Then copy and paste the three lines
of code within the samp.sd function into the pop.sd function. Update the three lines as necessary
so the pop.sd function will calculate the population standard deviation formula for a set of values.



(b) Run the entire chunk of code (from set.seed(1) through par(mfrow=c(1,1))). (i) What do you
notice about the average of the standard deviations produced using the samp.sd function compared
to the pop.sd function compared to the true population standard deviation? (ii) Why might we
prefer to use the sample.sd formulation when we have a sample of data and are interested in
estimating the population standard deviation? (You can compare the resulting histograms to help
you answer the question.)
<font color="red"> 
(i) The curve of the sample.sd is more like a bell curve. <br/>
(ii) There is less variance in the sample.sd function. 
</font>


```{r}
set.seed(1)
pop1<-rnorm(10000, 4, 2)
#Writing functions to calculate sample sd and population sd
samp.sd<-function(data){
n<-length(data)
sum.sq.devs<-sum((data-mean(data))^2)
av.dev<-sqrt(sum.sq.devs/(n-1))
return(av.dev)
}
pop.sd<-function(data){
n<-length(data)
sum.sq.devs<-sum((data-mean(data))^2)
av.dev<-sqrt(sum.sq.devs/(n))
return(av.dev)
}
#Simulation Section
#Building sampling distribution of pop standard deviation estimators
#estimator 1 is sd.sample and estimator 2 is sd.pop formulation
nsamples<-100000
sd.sample<-rep(0, nsamples)
sd.pop<-rep(0, nsamples)
for (i in 1:nsamples){
samp<-sample(pop1, 10, replace=TRUE) #taking a new sample of size 10 from population
sd.sample[i]<-samp.sd(samp) #calculating and storing sd using sample formula on new sample
sd.pop[i]<-pop.sd(samp) #calculating and storing sd using pop formula on new sample
}
#Displaying histograms of the 100000 standard deviations we calculated used
#the population and sample equations;
#adding true population standard deviation in red and
#mean of the simulated standard deviations in blue.
par(mfrow=c(1,2))
hist(sd.sample, freq=FALSE, xlim=c(0.3, 3.75)); #histogram of generated sample sds
abline(v=pop.sd(pop1), col="red"); abline(v=mean(sd.sample), col="blue")
hist(sd.pop, freq=FALSE, xlim=c(0.3, 3.75)); #histogram of generated pop sds
abline(v=pop.sd(pop1), col="red"); abline(v=mean(sd.pop), col="blue")
par(mfrow=c(1,1))
```