---
title: "Homework 6"
author: "Teryl Schmidt | tschmidt6@wisc.edu | 9072604920 | Discussion 312 | Grader: Chi-Shain Dai"
date: "11/1/2018"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Problem 1
An eel supply company has two large vats of adolescent eels (called ’elvers’). One vat contains American
eels, and the other contains European eels. An eel researcher suspects that American eels have a larger
number of scales than European eels. Let µAm be the population mean for American, and µEu be the
population mean for European. The species of eels can be considered independent. The researcher visits
the company and uses an ingenious method to get random samples from each vat (details omitted)
and counts the number of scales on each selected elver. Suppose the data from each sample supports
the assumption that both species scale counts are approximately normal. The sample statistics are
summarized below: <br/>
Type        |  Sample Size  |  Sample Mean  |  Sample Variance  | <br/>
American    |       8       |      220      |        21         | <br/>
European    |       5       |      204      |        23         | <br/>


(a) Explain why a two sample t test with an equal variance assumption is appropriate to use here.
List the null and alternative hypothesis.

<font color="red"> The two samples are independent, and normally distributed. The variances are close. So all the three assumptions are met for 2-sample t-test. <br/>
Ho: &mu;<sub>US</sub> - &mu;<sub>EU</sub> = 0 <br/>
Ha: &mu;<sub>US</sub> - &mu;<sub>EU</sub> > 0 <br/>
</font>

(b) At α = 0.1, find the rejection region in the scale of t and also in the scale of X¯Am − X¯Eu and use
them to make a reject or not reject decision based on the observed test statistic and observed
difference in means, respectively. Then make a conclusion in the context of the problem.

<font color="green">
The rejection region of the scaleof t is: t<sub>((8+5)-2 , 0.1)</sub> = 1.36 <br/>
Based on t = X¯Am − X¯Eu / s<sub>p</sub> * sqrt(1 / n<sub>Am</sub> + 1 / n<sub>Eu</sub>) <br/>
to reverse standardize we need to find the estimated standard error. <br/>
SE = sqrt(((7 * 21) + (4 * 23)) / 11) = 4.661252 <br/>
Reverse standardize get 1.36 * 4.661252 = 6.339303 <br/>
The rejection region is reject if X¯Am − X¯Eu > 6.339303 <br/>
<font color="red">
220 - 204 = 16 <br/>
16 > 6.339303 so reject the null hypothesis. <br/>
</font>
</font>

(c) Decide the same test based on the p-value compared to α = 0.1.

<font color="green"> 
SE = sqrt(((7 * 21) + (4 * 23)) / 11) = 4.661252
t<sub>obs</sub> = (220 - 204) / (4.66 * sqrt(1/8 + 1/5)) = 6.022716 <br/>
pvalue = P(t<sub>12</sub> > 6.022716) <br/>
pvalue < 0.0001 <br/>
<font color="red"> 
The null hypothesis is rejected. We conclude that the population mean for American is different from the population mean for European. <br/>
</font>
</font>

(d) Suppose we instead wanted to test:
H0 : µAm − µEu = 10
vs.
HA : µAm − µEu > 10
If the same data is used for both tests, would the p-value for this test be larger or smaller than
the p-value that would have been computed for the test in part (a)? Explain your answer. You do
not need to actually compute either p-value to answer this question.

<font color="red"> 
The pvalue would be bigger because 16 is closer to 10 than 0 is.
</font>

(e) Suppose instead we choose to perform a Welch’s t test. What does that mean for the assumptions
we are making? Perform a Welch’s T test for the same data. How much do the p values differ?

<font color="green"> 
t<sub>v</sub> = X¯Am − X¯Eu - 0 / sqrt(S<sup>2</sup><sub>Am</sub> / n<sub>Am</sub> + S<sup>2</sup><sub>Eu</sub> / n<sub>Eu</sub>) <br/>
= (220 - 204 - 0) / sqrt((21 / 8) + (23 / 5)) = 5.952523 <br/>
v = (220 - 204 - 0) / ((21 / 8) + (23 / 5))^2 / ((21 / 8)^2 + (23 / 5)^2)^2 = 241.1818
P(t<sub>241</sub> > 5.952523) < 0.0005 <br/>
= pvalue < 0.0005 <br/>
<font color="red"> 
We assume population and samples are independent, and that the data is normal. We do not assume equal variance. Since pvalue < 0.0005 there is strong evidence to suggest that the population mean for American is different from the population mean for European.
</font>
</font>

# Problem 2
Data on household vehicle miles of travel (VMT) are compiled annually by the Federal Highway
Administration. A researcher is interested in whether there is a difference in last year’s mean MVT for
midwestern and southern households. Independent random samples of 15 midwestern households and
14 southern households provided the following data on last year’s VMT, in thousands of miles: <br/>
Midwest : 16.2, 12.9, 17.3, 14.6, 18.6, 10.8, 11.2, 16.6, 16.6, 24.4, 20.3, 20.9, 9.6, 15.1, 18.3 <br/>
South : 22.2, 19.2, 9.3, 24.6, 20.2, 15.8, 18.0, 12.2, 20.1, 16.0, 17.5, 18.2, 22.8, 11.5 <br/>


(a) Graph the data as you see fit. Why did you choose the graph(s) you did and what does it (do
they) tell you? Also calculate summary statistics relevant to the research question.

```{r}
midwest = c(16.2, 12.9, 17.3, 14.6, 18.6, 10.8, 11.2, 16.6, 16.6, 24.4, 20.3, 20.9, 9.6, 15.1, 18.3)
south = c(22.2, 19.2, 9.3, 24.6, 20.2, 15.8, 18.0, 12.2, 20.1, 16.0, 17.5, 18.2, 22.8, 11.5)
name = c("Midwest", "South")
qqnorm(midwest, main = "Midwest QQ")
qqnorm(south, main = "South QQ")
hist(midwest, main = "Midwest Histogram")
hist(south, main = "South Histogram")
boxplot(midwest, south, horizontal = T, names = name)
mean(midwest)
mean(south)
sd(midwest)
sd(south)
```
<font color="red"> 
The QQ plots and histograms tells us if the data is skewed and normal, the boxplots shows how the data compares.
</font>

(b) Perform a two population t test for means at the 10% level assuming equal variance and justify
why the assumptions of the test are resonably met (or describe what assumptions we are assuming
are met). As part of this test, specify your hypotheses, calculate a p value and make a conclusion
in the context of the question. Compute the p value just using your calculator and t table. Then
find an exact p value by running the same test in R.

<font color="green">
Assumptions: Independence, Normality, Equal variance<br/>
Ho: &mu;<sub>M</sub> = &mu;<sub>S</sub> <br/>
Ha: &mu;<sub>M</sub> =/= &mu;<sub>S</sub> <br/>
S<sub>p</sub> = sqrt((((n<sub>M</sub> - 1) * s<sub>M</sub>^2) + ((n<sub>S</sub> - 1) * s<sub>S</sub>^2)) / ((n<sub>M</sub> + n<sub>S</sub> - 2)) <br/>
S<sub>p</sub> = sqrt((((15 - 1) * 4.06^2) + ((14 - 1) * 4.42^2)) / (15 + 14 - 2)) = 4.237153 <br/>
t = X¯<sub>M</sub> − X¯<sub>S</sub> / S<sub>p</sub> * sqrt(1 / n<sub>M</sub>  + 1 / n<sub>S</sub>) <br/>
t = 16.23 - 17.69 / 4.23715 * sqrt(1 / 15 + 1 / 14) <br/>
= -0.93 <br/>
+- t<sub>0.10 / 2</sub> <br/>
df = 15 + 14 - 2 = 27 <br/>
= +-t<sub>(27 , 0.05)</sub> = +- 1.703 <br/>
<font color="red"> 
The value of the test statistic t = -0.93 which does not fall in the rejection region. Thus we do not reject Ho. The test results are not significant at 10%.
</font>
</font>

(c) Construct a 90% confidence interval for the true difference in means (by hand and then again in
R). Describe how this confidence interval relates to your findings in part b.

<font color="green"> 
= t<sub>(27 , 0.05)</sub> = 1.703 <br/>
(X¯<sub>M</sub> − X¯<sub>S</sub>) +- t<sub>a / 2</sub> * S<sub>p</sub> * sqrt(1 / n<sub>M</sub>  + 1 / n<sub>S</sub>) <br/>
(16.23 - 17.69) +- (1.703) * (4.23715) * sqrt(1 / 15 + 1 / 14) <br/>
= (-4.14 , 1.22) <br/>
<font color="red"> 
For the two tailed test at 10% level the null hypothesis is not rejected and it is confirmed that the confidence interval  (-4.14 , 1.22) contains 0. <br/>
</font>
</font>
```{r}
t.test(midwest, south, var.equal=TRUE, mu=0, alternative = "two.sided", conf.level = 0.90)
```

# Problem 3
Several neurosurgeons wanted to determine whether a dynamic system (Z-plate) reduced operative time
relative to a static system (ALPS plate). The operative times, in minutes, for 14 dynamic replications
of the operataion and 6 static replications were obtained and are given below: <br/>
Dynamic : 370, 360, 510, 445, 295, 315, 490, 345, 450, 505, 335, 280, 325, 500 <br/>
Static : 430, 445, 455, 455, 490, 535 <br/>


(a) Graph the data as you see fit. Why did you choose the graph(s) you did and what does it (do
they) tell you? Also calculate summary statistics relevant to the research question.

```{r}
Dynamic = c(370, 360, 510, 445, 295, 315, 490, 345, 450, 505, 335, 280, 325, 500)
Static = c(430, 445, 455, 455, 490, 535)
qqnorm(Dynamic)
qqnorm(Static)
boxplot(Dynamic, Static, horizontal = T, names = c("Dynamic", "Static"))
```

<font color="red"> 
We assume SRS (simple random sample) and that the samples are independent. To check large sample size, we refer to the normal probability plots and boxplots. These graphs reveal no outliers and, keeping in mind that the nonpooled t-test is robust to moderate violations of normality, show that we can consider Assumption 3 satisfied.
The non-pooled t-test can be usedto carry out the hypothesis test.
</font>

(b) Perform a two population t test for means at the 5% level not assuming equal variance and justify
why the assumptions of the test are resonably met (or describe what assumptions we are assuming
are met). As part of this test, specify your hypotheses, calculate a p value and make a conclusion
in the context of the question. Compute the p value just using your calculator and t table. Then
find an exact p value by running the same test in R.

<font color="green"> 
Ho: &mu;<sub>D</sub> = &mu;<sub>S</sub> <br/>
Ha: &mu;<sub>D</sub> < &mu;<sub>S</sub> <br/>
t = X¯<sub>D</sub> − X¯<sub>S</sub> / sqrt(S<sup>2</sup><sub>D</sub> / n<sub>D</sub> + S<sup>2</sup><sub>S</sub> / n<sub>S</sub>) <br/>
t = (394.6 - 468.3) / sqrt((84.7^2 / 14) + (38.2^2 / 6)) = -2.681 <br/>
df = ((84.7^2 / 14) + (38.2^2 / 6))^2 / (((84.7^2 / 14)^2 / (14 - 1)) + ((38.2^2 / 6)^2 / (6 - 1))) ~ 17 <br/>
t<sub>(17 , -2.681)</sub> <br/>
0.005 < pvalue < 0.01 <br/>
If pvalue <= a, reject Ho, otherwise do not reject Ho. <br/>
<font color="red"> 
Because the pvalue is less than the specified significance level of 0.05, we reject Ho. The test results statistically significant at 5% level and provides strong evidence against the null hypothesis. <br/>
At the 5% significance level, the data provide sufficient evidence
to conclude that the mean operative time is less with the dynamic system than
with the static system.
</font>
</font>
```{r}
t.test(Dynamic, Static, var.equal=TRUE, mu=0, alternative = "less")
```

(c) Construct and interpret an 95% confidence interval for the true difference in means (by hand and
then again in R) applying the same assumptions as in part b. Describe how this confidence interval
relates to your findings in part b.

<font color="green"> 
= t<sub>(17 , 0.025)</sub> = 2.11 <br/>
(X¯<sub>M</sub> − X¯<sub>S</sub>) +- t<sub>a / 2</sub> * S<sub>p</sub> * sqrt(1 / n<sub>M</sub>  + 1 / n<sub>S</sub>) <br/>
(16.23 - 17.69) +- (2.11) * (4.23715) * sqrt(1 / 15 + 1 / 14) <br/>
= (-4.78 , 1.86) <br/>
<font color="red"> 
For the two tailed test at 5% level the null hypothesis is not rejected and it is confirmed that the confidence interval  (-4.78 , 1.86) contains 0. <br/>
</font>
</font>
```{r}
t.test(Dynamic, Static, var.equal=TRUE, mu=0, alternative = "two.sided", conf.level = 0.95)
```

# Problem 4
Two new mathematics learning techniques are being tested. Twenty students were randomly selected
from a population. nA = 9 of them were randomly assigned to use technique A, and nB = 11 of them
were randomly assigned to use technique B. Each student spent 30 minutes learning the technique to
which they were assigned, and then were asked to complete a task. The time to complete the task was
recorded, in seconds. A shorter time indicates better mastery of the task. The data are below: <br/>
TechniqueA : 23.1, 21.4, 31.6, 34.5, 21.9, 36.0, 30.2, 33.1, 39.5 <br/>
TechniqueB : 32.7, 36.8, 39.1, 37.3, 40.3, 46.8, 41.4, 53.0, 55.6, 54.1, 28.3 <br/>
We wish to test:
Ho : µA − µB = 0
vs.
HA : µA − µB =/= 0
using α = 0.05.


(a) Graph the data as you see fit. Why did you choose the graph(s) you did and what does it (do
they) tell you? Also calculate summary statistics relevant to the research question.

```{r}
A = c(23.1, 21.4, 31.6, 34.5, 21.9, 36.0, 30.2, 33.1, 39.5)
B = c(32.7, 36.8, 39.1, 37.3, 40.3, 46.8, 41.4, 53.0, 55.6, 54.1, 28.3)
barplot(A, ylim = c(0, 60), name = "Technique A")
barplot(B, ylim = c(0, 60), name = "Technique B")
qqnorm(A)
qqnorm(B)
boxplot(A,B, horizontal = T, names = c("A", "B"))
mean(A)
mean(B)
sd(A)
sd(B)
```
<font color="red"> 
A barplot is used to plot the several groups. It shows that technique B has a higher mean than A.
</font>


(b) Use the bootstrap to perform the test, using B = 10000 resamplings and set.seed(1). Display the
histogram of your generated tˆ values. Compute your tobs and a p-value. Make a reject or not
reject decision. Finally, state your conclusion in the context of the problem.

Bootstrapping Function:
```{r}
s.pooled=sqrt(((length(A)-1)*(sd(A))^2+(length(B)-1)*(sd(B))^2)/(length(A)-1+length(B)-1))
se.diff<-sqrt(s.pooled^2/10+s.pooled^2/12)
t.obs<-((mean(A)-mean(B))-0)/se.diff
# Here's one way to do the bootstrap for a difference of two means in R: 
# dat1 and dat2 are data from the two groups. nboot is the number of resamples.
#Notice obsdiff is computed as mean(dat1)-mean(dat2) - order matters!
boottwo = function(dat1, dat2, nboot) {
  bootstat = numeric(nboot)
  obsdiff = mean(dat1) - mean(dat2)
  n1 = length(dat1)
  n2 = length(dat2)
  for(i in 1:nboot) {
    samp1 = sample(dat1, size = n1, replace = T)
    samp2 = sample(dat2, size = n2, replace = T)
    bootmean1 = mean(samp1)
    bootmean2 = mean(samp2)
    bootvar1 = var(samp1)
    bootvar2 = var(samp2)
    bootstat[i] = ((bootmean1 - bootmean2) - obsdiff)/sqrt((bootvar1/n1) + (bootvar2/n2))
}
  return(bootstat)
}

B2 = 10000
set.seed(1)
boot = boottwo(A, B, B2) 
hist(boot)

(t.obs = (mean(A) - mean(B)) /
         sqrt(var(A) / length(A)  + var(B) / length(B)))

(m.low  = sum(boot < t.obs))
(m.high = sum(boot > t.obs))
2 * (p.val = m.low / B2)
```

<font color="red"> 
The pvalue < 0.05 so there is strong evidence to reject the null hypothesis, so the means of the two learning technniques are different.
</font>

(c) Use R to perform a two-group t test for means (i) assuming equal variance and then (ii) not
assuming equal variance (Welch’s T) and report the p values from each. Describe how the three p
values are related. Explain how the histogram in part (b) and the summary statistics in part (a)
hinted at this relationship.

```{r}
t.test(A, B, var.equal=FALSE, alternative="two.sided")
```

<font color="red"> 
The barplots show that almost all of the data in technique B is greater than in A, this shows that the means are going to be different.
</font>


