---
title: "Discussion2"
author: "Teryl Schmidt"
date: "9/18/2018"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
Method1 = c(0.32, 0.35, 0.37, 0.39, 0.42, 0.47, 0.51, 0.58, 0.60, 0.62, 0.65, 0.68, 0.75)
boxplot(Method1, ylim = c(0,1), horizontal = T)

Method2 = c(0.25, 0.40, 0.48, 0.55, 0.56, 0.58, 0.60, 0.65, 0.70, 0.76, 0.80, 0.91, 0.99)
boxplot(Method2, ylim = c(0,1), horizontal = T)

par(mfrow = c(1,1))

all_data = c(Method1, Method2)
label = rep(c("Method1", "Method2"), each = 13)

boxplot(all_data~label)
?boxplot

IQR(Method1)
IQR(Method2)
summary(Method1)
summary(Method2)

boxplot(Method1, Method2, horizontal = T)

```

Two dogs will be selected by lottery to receive the training. Assuming that each pair of dogs are equally
likely to receive the training calculate the following probabilities:
--
M 10
M 15
F 10
F 10
--

1,2
1,3
1,4
2,3
2,4
3,4



(a) P(A) where A is the event that the dogs are the same gender.
1,2 , 3,4

2 / 6 = 1 / 3


(b) P(B) where B is the event that the dogs are the same age.
1,3 1,4 3,4

3 / 6 = 1 / 2


(c) P(A or B)

P(A) = 1,2 3,4
P(B) = 1,3 1,4 3,4

P(A or B) = 1,2 1,3 1,4 3,4

4 / 6 = 2 / 3


(d) P(A and B)

P(A) = 1,2 3,4
P(B) = 1,3 1,4 3,4

P(A and B) = 3,4

1 / 6


(e) Are A and B mutually exclusive?

No, since P(A and B) = 3,4 which is not the empty set


(f) Are A and B independent?

iff P(A and B) = P(A) * P(B)
P(A) P(B) = 1/3 * 1/2 = 1/6
Yes because P(A and B) = P(A) * P(B)






